---
title: "Pressures and Resilience Figures"
author: "Jamie Montgomery"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(sf)
library(ggrepel)
library(paletteer)
library(gghighlight)
library(RColorBrewer)
library(patchwork)

# set the mazu data_edit share based on operating system
dir_M             <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
                       'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
                       'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]
dir_prep <- '~/github/ne-prep'
# shapefiles
# state land boundaries
ne_states  <- sf::st_read(dsn = paste0(path.expand(dir_prep),'/spatial/shapefiles'), layer = 'states', quiet = T)
rgns       <- sf::st_read(dsn = paste0(path.expand(dir_prep),'/spatial/shapefiles'), layer = 'ne_ohi_rgns', quiet = T) #need to use path.expand because readOGR does not read '~'
rgns_simp  <- sf::st_read(dsn = paste0(path.expand(dir_prep),'/spatial/shapefiles'),layer = 'ne_ohi_rgns_simp', quiet = T) #need to use path.expand because readOGR does not read '~'

ne_rgn <- sf::read_sf(dsn = paste0(path.expand(dir_prep),'/spatial/shapefiles'), layer = 'northeast_rgn', quiet = T) #whole region
rgn_data   <- data.frame(rgns) %>% 
  select(-geometry) %>% 
  mutate(area_km2 = as.numeric(levels(area_km2))[area_km2],
         rgn_name = as.character(rgn_name),
         state = ifelse(str_detect(rgn_name, "Massachusetts"), "Massachusetts", rgn_name))

scores <- read_csv("~/github/ne-scores/region/scores.csv") %>%
  left_join(rgn_data, by = c("region_id" = "rgn_id")) %>%
  select(-area_km2, -state, -state_abv, -state_name) %>%
  mutate(goal_name = case_when(
    goal == "BD" ~ "Biodiversity",
    goal == "CW" ~ "Clean Waters",
    goal == "ECO" ~ "LE: Economies",
    goal == "FIS" ~ "FP: Wild-Caught Fisheries",
    goal == "HAB" ~ "BD: Habitats",
    goal == "HS" ~ "Habitat Services",
    goal == "CP" ~ "HS: Coastal Protection",
    goal == "CS" ~ "HS: Carbon Storage",
    goal == "ICO" ~ "SOP: Iconic Species",
    goal == "Index" ~ "Ocean Health Index",
    goal == "LE" ~ "Livelihoods & Economies",
    goal == "LIV" ~ "LE: Livelihoods",
    goal == "LSP" ~ "SOP: Lasting Special Places",
    goal == "MAR" ~ "FP: Aquaculture",
    goal == "RAO" ~ "Resource Access Opportunities",
    goal == "SP" ~ "Sense Of Place",
    goal == "SPFIS" ~ "SOP: Fishing Engagement",
    goal == "TR" ~ "Tourism & Recreation",
    goal == "SPP" ~ "BD: Species",
    goal == "FP" ~ "Food Provision"
  ),
  rgn_name = ifelse(region_id == 0, "Northeast", rgn_name),
  year = as.numeric(year))
```


# Pressures

## Pressures by goal and region

Pressures are going up across all regions/goals.

```{r, fig.height = 7, fig.width = 9}
pressures <- scores %>%
  filter(dimension == "pressures",
         !is.na(score)) 

ggplot(pressures, aes(x = year, y = score, color = rgn_name)) +
  geom_line() +
  facet_wrap(~goal_name) +
  theme_minimal()

```


Look at pressures individually

```{r}
#list all pressures from ne-scores/region/layers

pressure_list <- list.files("~/github/ne-scores/region/layers", full.names = T)[
  str_detect(list.files("~/github/ne-scores/region/layers", full.names = T), "prs_|cc_|hd_|sp_a|sp_g")]

#read each one in and combine into a dataframe
out <- data.frame()
for(i in 1:length(pressure_list)){
  print(i)
  name <- substr(pressure_list[i],49, nchar(pressure_list[i])-4)
  
  #issue with pathogens
  if(name == "prs_pathogens"){
    df <- read_csv(pressure_list[i]) %>%
    mutate(pressure = name) %>%
    select(year, rgn_id, score = perc_closed, pressure)
    }
  
  if(name == "prs_hardening"){
    df <- read_csv(pressure_list[i]) %>%
    mutate(pressure = name) %>%
    select(year, rgn_id, score = percent_hard, pressure) 
  }
  
  if(!name %in% c("prs_pathogens", "prs_hardening")){
  df <- read_csv(pressure_list[i]) %>%
    mutate(pressure = name) %>%
    select(year, rgn_id, contains("rgn_name"), contains("score"), pressure)
  }
  
  out <- bind_rows(df, out)
}

#clean up
pressure_df <- out %>%
  mutate(score = ifelse(is.na(pressure_score), score, pressure_score)) %>%
  left_join(rgn_data, by = "rgn_id") %>%
  select(rgn_id, rgn_name = rgn_name.y, score, pressure, year) %>%
  mutate(rgn_name = ifelse(rgn_id == 12, "Northeast", rgn_name),
         pressure_full = case_when(
           pressure == "cc_acid" ~ "Ocean acidification",
           pressure == "cc_slr" ~ "Sea level rise",
           pressure == "cc_sst" ~ "Sea surface temperature",
           pressure == "prs_wqi" ~ "Water quality",
           pressure == "prs_sqi" ~ "Sediment quality",
           pressure == "prs_trash" ~ "Coastal trash",
           pressure == "prs_pathogens" ~ "Pathogens",
           pressure == "hd_intertidal" ~ "Intertidal habitat destruction",
           pressure == "hd_subtidal" ~ "Subtidal habitat destruction",
           pressure == "prs_fishing" ~ "Fishing",
           pressure == "prs_social" ~ "Social",
           pressure == "prs_aqua_hab" ~ "Habitat destruction from aquaculture",
           pressure == "prs_hardening" ~ "Coastal hardening",
           pressure == "prs_disease" ~ "Disease from aquaculture",
           pressure == "sp_alien" ~ "Invasive species",
           pressure == "sp_genetic" ~ "Genetic escapes")) %>%
  filter(year > 2004)
```

Some of these layers have values for all years while others do not. This is taken care of in the toolbox using the `scenario_years.csv` file in ne-scores but I need to fix it here which can be done using `fill()`

```{r}
fill_prs_df <- pressure_df %>%
  group_by(rgn_id, rgn_name, pressure, pressure_full) %>%
  complete(year = 2005:2017) %>%
  ungroup() %>%
  group_by(pressure, rgn_id, rgn_name, pressure_full) %>%
  fill(score, .direction = "downup")
```



```{r, fig.width = 10, fig.height = 4.5}
ggplot() +
  geom_line(data = fill_prs_df, aes(x = year, y = score, color = pressure_full), lty=2) +
  geom_line(data = pressure_df, aes(x = year, y = score, color = pressure_full)) +
  theme_minimal() +
  facet_wrap(~rgn_name) +
  scale_x_continuous(breaks=seq(2005, 2017, 3)) +
  labs(x = "",
       y = "Score",
       color = "Pressure")
```
Split into groups

```{r,  fig.width = 8, fig.height = 4.5}
#climate
climate <- ggplot() +
  geom_line(data = filter(fill_prs_df, str_detect(pressure, "cc_")), aes(x = year, y = score, color = pressure_full), lty=2) +
  geom_line(data = filter(pressure_df, str_detect(pressure, "cc_")), aes(x = year, y = score, color = pressure_full)) +
  theme_minimal() +
  facet_wrap(~rgn_name) +
  scale_x_continuous(breaks=seq(2005, 2017, 3)) +
  labs(x = "",
       y = "Score",
       title = "Climate pressures",
       color = "") +
  theme(axis.text.x = element_text(size = 7),
        legend.position = "bottom")

climate

ggsave("../figs/supp_fig_climate_pressures_over_time.jpg")
```
Pollution
```{r,  fig.width = 6, fig.height = 4.5}
#pollution
pollution <- ggplot() +
  geom_line(data = filter(fill_prs_df, str_detect(pressure, "sqi|wqi|trash|path") & !is.na(score)), aes(x = year, y = score, color = pressure_full), lty=2) +
  geom_line(data = filter(pressure_df, str_detect(pressure, "sqi|wqi|trash|path") & !is.na(score)), aes(x = year, y = score, color = pressure_full)) +
  theme_minimal() +
  facet_wrap(~rgn_name) +
  scale_x_continuous(breaks=seq(2005, 2017, 3)) +
  ylim(0,1) +
  labs(x = "",
       y = "Score",
       title = "Pollution pressures",
       color = "") +
  theme(axis.text.x = element_text(size = 7),
        legend.position = "bottom")

pollution

ggsave("../figs/supp_fig_pollution_pressures_over_time.jpg")
```

Habitat destruction

```{r,  fig.width = 8, fig.height = 4.5}

hab_dest <- ggplot() +
  geom_line(data = filter(fill_prs_df, str_detect(pressure, "hd|aqua|hard") & !is.na(score)), aes(x = year, y = score, color = pressure_full), lty=2) +
  geom_line(data = filter(pressure_df, str_detect(pressure, "hd|aqua|hard") & !is.na(score)), aes(x = year, y = score, color = pressure_full)) +
  theme_minimal() +
  facet_wrap(~rgn_name) +
  scale_x_continuous(breaks=seq(2005, 2017, 3)) +
  ylim(0,1) +
  labs(x = "",
       y = "Score",
       title = "Habitat destruction pressures",
       color = "") +
  theme(axis.text.x = element_text(size = 7),
        legend.position = "bottom")

hab_dest

ggsave("../figs/supp_fig_hab_dest_pressures_over_time.jpg")
```

Others

```{r,  fig.width = 8, fig.height = 4.5}
others <- ggplot() +
  geom_line(data = filter(fill_prs_df, str_detect(pressure, "fish|social|dis|sp_") & !is.na(score)), aes(x = year, y = score, color = pressure_full), lty=2) +
  geom_line(data = filter(pressure_df, str_detect(pressure, "fish|social|dis|sp_") & !is.na(score)), aes(x = year, y = score, color = pressure_full)) +
  theme_minimal() +
  facet_wrap(~rgn_name) +
  scale_x_continuous(breaks=seq(2005, 2017, 3)) +
  ylim(0,1) +
  labs(x = "",
       y = "Score",
       title = "Other pressures",
       color = "") +
  theme(axis.text.x = element_text(size = 7),
        legend.position = "bottom")

others
ggsave("../figs/supp_fig_other_pressures_over_time.jpg")
```

# Resilience

```{r}

#list all pressures from ne-scores/region/layers

res_list <- list.files("~/github/ne-scores/region/layers", full.names = T)[
  str_detect(list.files("~/github/ne-scores/region/layers", full.names = T), "/res_")]

#read each one in and combine into a dataframe
out <- data.frame()
for(i in 1:length(res_list)){
  print(i)
  name <- substr(res_list[i],49, nchar(res_list[i])-4)
  
  df <- read_csv(res_list[i]) %>%
    mutate(resilience = name) %>%
    select(year, rgn_id, contains("rgn_name"), contains("score"), resilience)
  
  out <- bind_rows(df, out)
}

#clean up
res_df <- out %>%
  mutate(score = ifelse(is.na(resilience_score), score, resilience_score)) %>%
  left_join(rgn_data, by = "rgn_id") %>%
  select(rgn_id, rgn_name = rgn_name.y, score, resilience, year) %>%
  mutate(rgn_name = ifelse(rgn_id == 12, "Northeast", rgn_name),
         res_full = case_when(
           resilience == "res_alien" ~ "Alien species",
           resilience ==  "res_climate_change" ~ "Climate change",
           resilience ==  "res_eco_integrity" ~ "Ecological integrity",
           resilience ==  "res_fishing" ~ "Fishing",
           resilience ==  "res_hab_destruction" ~ "Habitat destruction",
           resilience ==  "res_social" ~ "Social",
           resilience ==  "res_water_pollution" ~ "Water pollution")) %>%
  filter(year > 2004)
```

Some of these layers have values for all years while others do not. This is taken care of in the toolbox using the `scenario_years.csv` file in ne-scores but I need to fix it here which can be done using `fill()`

```{r}
fill_res_df <- res_df %>%
  group_by(rgn_id, rgn_name, resilience, res_full) %>%
  complete(year = 2005:2017) %>%
  ungroup() %>%
  group_by(resilience, rgn_id, rgn_name, res_full) %>%
  fill(score, .direction = "downup")
```

```{r, fig.width = 8, fig.height = 4.5}
ggplot() +
  geom_line(data = fill_res_df, aes(x = year, y = score, color = res_full), lty=2) +
  geom_line(data = res_df, aes(x = year, y = score, color = res_full)) +
  theme_minimal() +
  facet_wrap(~rgn_name) +
  scale_x_continuous(breaks=seq(2005, 2017, 3)) +
  labs(x = "",
       y = "Score",
       color = "") +
  ylim(0,1)+
  theme(axis.text.x = element_text(size = 7),
        legend.position = "bottom")

ggsave("../figs/resilience_scores_by_region.jpg")
```